NMRlipids IV: Headgroup \& glycerol backbone structures, and cation binding in bilayers with PE and PG lipids

### 2018-09-20 Manuscript/manuscriptPGPE.tex
Add todos to PGPE manuscript

@@

\subsection{PE headgroup and glycerol backbone}
Order parameters from Slipids simulations 
and experiments for DPPE are shown in Fig. \ref{HGorderParametersPE}.
Glycerol backbone order parameters in Slipids are off from experiments,
as already observed previously for PC lipids \cite{botan15}.
Order parameter signs for PE are not experimentally measured yet.
For headgroup the signs are set to give best agreement with simulations
and for glycerol to be consistent with experimental signs for PC.
Order parameter for $\beta$ carbon shows apparent agreement
with experiments. However, the sign of beta order parameter is positive,
in contrast to PC where negative sign was measured. Thus, the the beta order
parameter agrees with experiment with the assumption that its sign is opposite
than for PC. This is yet to be confirmed by experiments.
Order parameter for $\alpha$ carbon is too close to zero, even if the sign
would be correct.

@@

@@

\section{Simulated systems}
\begin{table*}[!p]
%\begin{sidewaystable*}[!p]
\centering
\caption{List of MD simulations. The salt concentrations calculated as 
   [salt]=N$_{\rm c} \times$[water]\,/\,N$_{\rm w}$, where [water]\,=\,55.5~M.
%   these correspond the concentrations reported in the experiments by Akutsu et al.~\cite{akutsu81}.
%   The lipid force fields named as in our previous work~\cite{botan15}.
}\label{IONsystems}
\begin{minipage}[t]{\textwidth}
\begin{tabular}{l c c r r r r r r c c}
  %\hline
  % some footnotes are not visible in typeset-MS (pdf)
 lipid/counter-ions & force field for lipids / ions & NaCl (mM) & CaCl$_2$\,(mM) &  \footnote{Number of lipid molecules with largest mole fraction}N$_{\rm l}$   &  \footnote{Number of water molecules}N$_{\rm w}$   & \footnote{Number of additional cations}N$_{\rm c}$  & \footnote{Simulation temperature}T (K)  & \footnote{Total simulation time}t$_{{\rm sim}}$(ns) & \footnote{Time used for analysis}t$_{{\rm anal}}$ (ns) &   \footnote{Reference for simulation files}files\\
  \hline
  DPPE  & Slipids \cite{jambeck12b} &0 & 0        & 288 			   		& 9386 & 0  & 336  & 200 & 100 & \cite{slipidsDPPEfiles}  \\
    \hline
    DOPS/Na$^+$  & CHARMM36 \cite{??}  \todoi{Correct citation for CHARMM DOPS}      &0 & 0        & 128 			   		& 4480 & 0  & 303  & 500 & 100 & \cite{??}
    \todoi{By Piggot: http://nmrlipids.blogspot.com/2017/03/nmrlipids-iv-headgroup-glycerol.html?showComment=1491425687561\#c4932902612512697301. We need to decide the switching version or discuss this somehow.}  \\
    DOPS/Na$^+$  & CHARMM36ua \cite{??} \todoi{Correct citation for CHARMMua DOPS}        &0 & 0        & 128 			   		& 4480 & 0  & 303  & 500 & 100 & \cite{??}
    \todoi{Delivered by Piggot. We need to decide the switching version or discuss this somehow. Data to be uploaded in Zenodo?}  \\
    DOPS/Na$^+$  & Slipids \cite{jambeck13}        &0 & 0        & 128 			   		& 4480 & 0  & 303  & 500 & 100 & \cite{??}
    \todoi{Delivered by Piggot. We need to decide the cut-off version or discuss this somehow. Data to be uploaded in Zenodo?}  \\
    DOPS/Na$^+$  & Slipids \cite{jambeck13}        &0 & 0        & 288 			   		& 11232 & 0  & 303  & 200 & 100 & \cite{slipidsDOPSfiles} \\
    DOPS/Na$^+$  & Berger \cite{mukhopadhyay04}        &0 & 0        & 128 			   		& 4480 & 0  & 303  & 500 & 100 & \cite{??}
    \todoi{Delivered by Piggot. Data to be uploaded in Zenodo?}  \\
    DOPS/Na$^+$  & GROMOS-CKP \cite{??} \todoi{Correct citation(s) for CKP.}       &0 & 0        & 128 			   		& 4480 & 0  & 303  & 500 & 100 & \cite{??}
    \todoi{Delivered by Piggot. We need to decide between RF and PME or discuss this somehow. Data to be uploaded in Zenodo?}  \\
    \hline
    POPS/Na$^+$  & CHARMM36 \cite{??} \todoi{Correct citation for CHARMM POPS}        &0 & 0        & 128 			   		& 4480 & 0  & 298  & 500 & 100 & \cite{??}
    \todoi{Delivered by Piggot. We need to decide the switching version or discuss this somehow. Data to be uploaded in Zenodo?}  \\
    POPS/Na$^+$  & CHARMM36ua \cite{??} \todoi{Correct citation for CHARMMua DOPS}        &0 & 0        & 128 			   		& 4480 & 0  & 298  & 500 & 100 & \cite{??}
    \todoi{Delivered by Piggot. We need to decide the switching version or discuss this somehow. Data to be uploaded in Zenodo?}  \\
    POPS/Na$^+$  & Slipids \cite{jambeck13}        &0 & 0        & 128 			   		& 4480 & 0  & 298  & 500 & 100 & \cite{??}
    \todoi{Delivered by Piggot. We need to decide the cut-off version or discuss this somehow. Data to be uploaded in Zenodo?}  \\
    \hline
    POPC:POPS (5:1)/Na$^+$  & CHARMM36 \cite{klauda10,??} \todoi{Correct citation for CHARMM POPS}        &0 & 0        & ? 			   		& ? & 0  & ?  & ? & ? & \cite{??}
    \todoi{Details to be filled and data to be uploaded in Zenodo by Ollila.}  \\
    \hline
    POPG/Na$^+$  & CHARMM36 \cite{??} \todoi{Correct citation for CHARMM POPG}        &0 & 0        & ? 			   		& ? & 0  & ?  & ? & ? & \cite{??}
    \todoi{Details to be filled and data to be uploaded in Zenodo by Ollila.}  \\
    POPG/Na$^+$  & Slipids \cite{jambeck13}     &0 & 0        & 288 			   		& 10664 & 0  & 298  & 250 & 100 & \cite{slipidsPOPGfiles} \\
    \hline
    DPPG/Na$^+$  & Slipids \cite{jambeck13}     &0 & 0        & 288 			   		& 11232 & 0  & 314  & 200 & 100 & \cite{slipidsDPPGfiles} \\
    DPPG/Na$^+$  & Slipids \cite{jambeck13}     &0 & 0        & 288 			   		& 11232 & 0  & 298  & 400 & 100 & \cite{slipidsDPPGfilesT298K} \\
\end{tabular}
\end{minipage}
%\end{sidewaystable*} 
\end{table*}

>>>>

\section{Methods}
\section{Molecular dynamics simulations}
Molecular dynamics simulation data was collected using open collaboration \cite{botan15}.
The simulations are listed in table \ref{systems}.
\begin{table*}[htb]
  %\begin{sidewaystable*}[!p]
  \centering
  \caption{List of MD simulations. The salt concentrations calculated as 
    [salt]=N$_{\rm c} \times$[water]\,/\,N$_{\rm w}$, where [water]\,=\,55.5~M.
    %   these correspond the concentrations reported in the experiments by Akutsu et al.~\cite{akutsu81}.
    %   The lipid force fields named as in our previous work~\cite{botan15}.
  }\label{systems}
  \begin{minipage}[t]{\textwidth}
    \begin{tabular}{l c c r r r r r r c c}
      %\hline
      % some footnotes are not visible in typeset-MS (pdf)
      lipid/counter-ions & force field for lipids / ions & NaCl (mM) & CaCl$_2$\,(mM) &  \footnote{Number of lipid molecules with largest mole fraction}N$_{\rm l}$   &  \footnote{Number of water molecules}N$_{\rm w}$   & \footnote{Number of additional cations}N$_{\rm c}$  & \footnote{Simulation temperature}T (K)  & \footnote{Total simulation time}t$_{{\rm sim}}$(ns) & \footnote{Time used for analysis}t$_{{\rm anal}}$ (ns) &   \footnote{Reference for simulation files}files\\
      \hline
      DPPE  & Slipids \cite{jambeck12b} &0 & 0        & 288 			   		& 9386 & 0  & 336  & 200 & 100 & \cite{slipidsDPPEfiles}  \\
      \hline
      POPG/Na$^+$  & CHARMM36 \cite{??} \todoi{Correct citation for CHARMM POPG}        &0 & 0        & ? 			   		& ? & 0  & ?  & ? & ? & \cite{??}
      \todoi{Details to be filled and data to be uploaded in Zenodo by Ollila.}  \\
      POPG/Na$^+$  & Slipids \cite{jambeck13}     &0 & 0        & 288 			   		& 10664 & 0  & 298  & 250 & 100 & \cite{slipidsPOPGfiles} \\
      \hline
      DPPG/Na$^+$  & Slipids \cite{jambeck13}     &0 & 0        & 288 			   		& 11232 & 0  & 314  & 200 & 100 & \cite{slipidsDPPGfiles} \\
      DPPG/Na$^+$  & Slipids \cite{jambeck13}     &0 & 0        & 288 			   		& 11232 & 0  & 298  & 400 & 100 & \cite{slipidsDPPGfilesT298K} \\
    \end{tabular}
  \end{minipage}
  %\end{sidewaystable*} 
\end{table*}

@@

\subsection{PE headgroup and glycerol backbone}
\todo{Add results from this discussion: http://nmrlipids.blogspot.com/2017/03/nmrlipids-iv-headgroup-glycerol.html?showComment=1502290977005#c4964787551006026647}
Order parameters from Slipids simulations 
and experiments for DPPE are shown in Fig. \ref{HGorderParametersPE}.
Glycerol backbone order parameters in Slipids are off from experiments,
as already observed previously for PC lipids \cite{botan15}.
Order parameter signs for PE are not experimentally measured yet.
For headgroup the signs are set to give best agreement with simulations
and for glycerol to be consistent with experimental signs for PC.
Order parameter for $\beta$ carbon shows apparent agreement
with experiments. However, the sign of beta order parameter is positive,
in contrast to PC where negative sign was measured. Thus, the the beta order
parameter agrees with experiment with the assumption that its sign is opposite
than for PC. This is yet to be confirmed by experiments.
Order parameter for $\alpha$ carbon is too close to zero, even if the sign
would be correct.

@@

\todo{There are mass density profiles available from CHARMM simulations, preliminary figures should be made:
http://nmrlipids.blogspot.com/2017/03/nmrlipids-iv-headgroup-glycerol.html?showComment=1501446779291#c7311317114567929503}

@@
